---
title: "p8105_mtp_zf2211"
author: "Francis"
date: "10/23/2018"
output: github_document
---

```{r}
#load packages
library(tidyverse)
library(ggridges)
library(gridExtra)
```

```{r}
#load data
atvy <- read.csv("./data/p8105_mtp_data.csv") %>% 
  janitor::clean_names()
nrow(atvy)
ncol(atvy)
```
```{r}
# Define normal order of weekdays
atvy$day <- factor(atvy$day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
atvy$week <- as.factor(atvy$week)
#tidy data
atvy_tidy <- 
  atvy %>%
  gather(key = minute, value = activity, activity_1:activity_1440) %>% 
  arrange(week, day)
#extract minute number
atvy_tidy$minute <- as.numeric(substr(atvy_tidy$minute, 10, 13))
# add a column to show hour
atvy_tidy <- 
  atvy_tidy %>% 
  mutate(hour = ceiling(minute / 60)) 
```
From the Data, one minute epoch are common, which means only when the activity in cells == 1 it is normal. When the activity counts are too high, the man would have been physical execrising. The dataset after tidy have `r ncol(atvy_tidy)` variables: `week` `day` `minute` `activity` `hour`and `r nrow(atvy_tidy)` observations. 

```{r}
ggplot(atvy_tidy, aes(x = day, y = activity)) +
  labs (title = "fig1. Activity Counts in Everyday in a Week (Scatterplot)") +
  geom_point()
```

From this scatterplot, activity counts(ACs) are relatively lower on Monday and Tuesday. In this year, outliers are: 1(>10000) on Sunday, 2(~7500) on Monday, 3(~7000, ~8000, ~9000) on Wednesday, 2(~9000) on Thursday, 1(~8000) on Saturday. Which is to say, in a week, the activity decrease from Sunday to Monday, then increase from Tuesday to Wednesday and keep to Saturday.


```{r fig.height = 5, fig.width = 9}
ggplot(atvy_tidy, aes(x = week, y = activity)) +
  labs (title = "fig2. Activity Counts in Each Week (Scatterplot)") +
  geom_point()
```
We can see ACs are lowered to 0 on 3rd week. Outliers are on week: 1st(>5000), 4th(~7000), 7th(~8000), 11th(~7500), 14th(~9000), 22nd(~7000), 30th(~4500), 33rd(~7500`*`2), 34th(~5000`*`2), 35th(~5000), 36th(~9000), 38th(~6000), 39th(~9500), 41st(~9000`*`2), 42nd(~6000`*`2), 43rd(~7000), 45th(~7000), 46th(~9000*2), 47(~8000). 
```{r}
#create total activity variable
mutate(atvy, total = sum(activity_1:activity_1440))

```

Here, variable `total` is the total activity variable of daily activity.

```{r fig.height = 5, fig.width = 10}
ggplot(atvy_tidy, aes(x = activity, y = day)) + 
  labs (title = "fig3. Activity Counts in Everyday in a Week (Ridgeplot)") +
  geom_density_ridges()
```

Over the everyday in a week, most of the activity are around 1. On Tuesday, Wednesday and Thursday, There are small aggregation other than 1. Which means on these three days, the ACs are increased to some extent. That is to say, activity increases from Monday to Tuesday, and decreases from Thursday to Friday.


```{r fig.height = 10, fig.width = 7}
ggplot(atvy_tidy, aes(x = activity, y = week)) + 
  labs(title = "fig4. Activity Counts in Each Week (Ridges Gradient plot)") +
  geom_density_ridges_gradient()
```

In each week, ACs are aggregated around 1. However, from 1 - 47, the peak of 1 is decreasing, in other words, the higher ACs are increasing. Especially, in 16th and 35th week, there are obvious peak other than at 1, which shows ACs increasing. Which is to say, the participant became more active over time.


```{r fig.height = 20, fig.width = 10}
a_week1 <-  atvy_tidy %>% filter(week == 1:10)
a_week2 <-  atvy_tidy %>% filter(week == 11:20)
a_week3 <-  atvy_tidy %>% filter(week == 21:30)
a_week4 <-  atvy_tidy %>% filter(week == 31:40)
a_week5 <-  atvy_tidy %>% filter(week == 41:47)
w1 = ggplot(a_week1, aes(x = hour, y = activity, color = week)) + 
  labs (title = "fig5. Daily Activity Count Trend") +
  stat_smooth(se = FALSE)
w2 = ggplot(a_week2, aes(x = hour, y = activity, color = week)) + 
  stat_smooth(se = FALSE)
w3 = ggplot(a_week3, aes(x = hour, y = activity, color = week)) + 
  stat_smooth(se = FALSE)
w4 = ggplot(a_week4, aes(x = hour, y = activity, color = week)) + 
  stat_smooth(se = FALSE)
w5 = ggplot(a_week5, aes(x = hour, y = activity, color = week)) + 
  stat_smooth(se = FALSE)
grid.arrange(w1, w2, w3, w4, w5, nrow = 5)
```

We can see in a whole day(averaged in a week), AC rises in the morning, and grow to a high plateau and extend for about 10 hours and goes down to near 1. Which overlaps the sleep-awake rythm. There is a flat curve in 12th week and even a level line in 3rd week. Maybe these are fault data like the accelerometer defect.




```{r}
#count word numbers
wordcountaddin::text_stats("p8105_mtp_zf2211.Rmd")
```

